|

### 2023-06-02 manuscript/paper.Rmd
Added plot of fossil sampel size per time period.

We used a third approach to infer fossil sex and study the evolution of sexual dimorphism by taking advantage of the fact that G. doryssus is part of the extant threespine stickleback species complex (Gastersteus aculeatus).  Along with the fossil data collected for this study, we also have data from extant populations of the Threespine Stickleback, Gasterosteus aculeatus, measuring the same traits for individuals whose sex could be determined directly.
We then combined these sets of data and treat the sex of the fossil specimens as missing and  used multiple imputation (@little2002statistical) to impute the sex of the fossils many times.  Imputations were performed using the MICE (@MICE) algorithm and implemented in R (@R2022language) with M = 100 imputations performed.

@@

# Data {#sec:data}
```{r}
stickle <- read_csv('stickle_revamped.csv')

@@

#fossil n
fortab <- data.frame(time = 1:18, count = as.character(as.numeric((table(times_vector)))))
#knitr::kable(fortab)
#print(xtable(fortab),include.rownames = FALSE)
```

@@

The fossil data consists of `r sum(is.na(stick$gender))` specimens with unknown sex over 18 time periods spaced  ~1000 years apart. Table 2 shows the sample size at each of the 18 time periods.  There are at least 22 specimens at each time period with a high of 67 specimens in period 7.

@@

<!-- % latex table generated in R 4.2.2 by xtable 1.8-4 package -->
<!-- % Mon May  1 11:20:28 2023 -->
\begin{table}[ht]
\centering
\caption{The number of stickleback fossils at each time point.  Sample size at each time point ranges from a low of 22 to a high of 67.}
\begin{tabular}{rl}
  \hline
time & count \\ 
  \hline
  1 & 43 \\ 
    2 & 41 \\ 
    3 & 51 \\ 
    4 & 41 \\ 
    5 & 46 \\ 
    6 & 48 \\ 
    7 & 67 \\ 
    8 & 55 \\ 
    9 & 42 \\ 
   10 & 33 \\ 
   11 & 37 \\ 
   12 & 22 \\ 
   13 & 41 \\ 
   14 & 43 \\ 
   15 & 46 \\ 
   16 & 47 \\ 
   17 & 56 \\ 
   18 & 55 \\ 
   \hline
\end{tabular}
\label{tabn}

\end{table}

>>>>

We used a third approach to infer fossil sex and study the evolution of sexual dimorphism by taking advantage of the fact that G. doryssus is part of the extant threespine stickleback species complex (Gastersteus aculeatus).  Along with the fossil data collected for this study, we also have data from extant populations of the Threespine Stickleback, Gasterosteus aculeatus, measuring the same traits for individuals whose sex could be determined directly.

We then combined these sets of data and treat the sex of the fossil specimens as missing and  used multiple imputation (@little2002statistical) to impute the sex of the fossils many times.  Imputations were performed using the MICE (@MICE) algorithm and implemented in R (@R2022language) with M = 100 imputations performed.

@@

# Data {#sec:data}
```{r}
library(tidyverse)
stickle <- read_csv('stickle_revamped.csv')

@@

#fossil n
#fortab <- data.frame(time = 1:18, count = as.character(as.numeric((table(times_vector)))))
fortab <- data.frame(time = 1:18, count = as.numeric((table(times_vector))))
#knitr::kable(fortab)
#print(xtable(fortab),include.rownames = FALSE)
```

@@

The fossil data consists of `r sum(is.na(stick$gender))` specimens with unknown sex over 18 time periods spaced  ~1000 years apart. Figure 1 shows the sample size at each of the 18 time periods.  There are at least 22 specimens at each time period with a high of 67 specimens in period 7.

```{r fig.cap = "Fossil sampel size for each time period"}
ggplot(aes(x = time, y = count), data = fortab) + geom_col() + xlab("Time Period")  + scale_x_continuous(breaks = 1:18, labels = 1:18)
```

@@

<!-- I removed the table and replaced it with a figure -->
<!-- % latex table generated in R 4.2.2 by xtable 1.8-4 package -->
<!-- % Mon May  1 11:20:28 2023 -->
<!-- \begin{table}[ht] -->
<!-- \centering -->
<!-- \caption{The number of stickleback fossils at each time point.  Sample size at each time point ranges from a low of 22 to a high of 67.} -->
<!-- \begin{tabular}{rl} -->
<!--   \hline -->
<!-- time & count \\  -->
<!--   \hline -->
<!--   1 & 43 \\  -->
<!--     2 & 41 \\  -->
<!--     3 & 51 \\  -->
<!--     4 & 41 \\  -->
<!--     5 & 46 \\  -->
<!--     6 & 48 \\  -->
<!--     7 & 67 \\  -->
<!--     8 & 55 \\  -->
<!--     9 & 42 \\  -->
<!--    10 & 33 \\  -->
<!--    11 & 37 \\  -->
<!--    12 & 22 \\  -->
<!--    13 & 41 \\  -->
<!--    14 & 43 \\  -->
<!--    15 & 46 \\  -->
<!--    16 & 47 \\  -->
<!--    17 & 56 \\  -->
<!--    18 & 55 \\  -->
<!--    \hline -->
<!-- \end{tabular} -->
<!-- \label{tabn} -->

<!-- \end{table} -->
